---
title: "Unusually Devastating Tornadoes"
author: "Tyler Fricker"
date: "6/10/2019"
output: html_document
---
Load packages.
```{r}
library(ggplot2)
library(dplyr)
library(lubridate)
library(sf)
```

Import `SocialCorrelates.shp` containing the casualty-producing tornadoes 1995-2016 and estimates of socioeconomic and demographic data at the tornado level. Add `Date` and `hr` columns to the data frame.
```{r}
unzip("SocialCorrelates.zip")
Tor.sfdf <- st_read(dsn = "SocialCorrelates", 
                    layer = "SocialCorrelates", 
                    stringsAsFactors = FALSE) %>%
  mutate(Date = as.Date(date),
         hr = hour(DateTim))
```

### Data Analysis

Preliminary analysis
```{r}
as.data.frame(Tor.sfdf) %>%
  filter(cas > 0) %>%
  summarize(meanCas = mean(cas),
            varCas = var(cas),
            ratio = varCas/meanCas)
```

Given the large disparity between the mean and variance of casualties the casualty counts are over-dispersed relative to a Poisson distribution. This suggests that a negative binomial (or log normal) distribution may be a reasonable likelihood model.

The number of tornadoes and the number of casualties conditional on at least one mobile home. 
```{r}
as.data.frame(Tor.sfdf) %>%
  group_by(MoblHms >= 1) %>%
  summarize(nC = sum(cas),
            nT = n())
```
About 13% of all tornado casualties occurred in areas estimated to have at least one mobile home.

Look at trends in annual tornado casualties and casualty-producing tornadoes.
```{r}
dfY <- Tor.sfdf %>%
  group_by(yr) %>%
  summarize(nCas = sum(cas),
            ncT = n())
  
ggplot(dfY, aes(x = yr, y = nCas)) +
  geom_point() +
  geom_smooth(method = lm) +
  scale_x_continuous(breaks = seq(1995, 2016, 2)) +
  xlab("Year") + ylab("Number of Tornado Casualties") +
  theme_minimal()

ggplot(dfY, aes(x = yr, y = ncT)) +
  geom_point() +
  geom_smooth(method = lm) +
  scale_x_continuous(breaks = seq(1995, 2016, 2)) +
  scale_y_continuous(limits = c(0, NA)) +
  xlab("Year") + ylab("Number of Casualty-Producing Tornadoes") +
  theme_minimal()
```

Look at trends in monthly tornado casualties. Points +/- 2 s.e.
```{r}
dfM <- Tor.sfdf %>%
  group_by(mo) %>%
  summarize(nT = n(),
            nCas = sum(cas),
            perTorCas = nCas/nT) %>%
  mutate(MonthF = factor(month.abb[mo], levels = month.abb),
         ci = qt(0.975, df = nT - 1) * sd(perTorCas)/sqrt(nT))

A = ggplot(dfM, aes(x = mo, y = perTorCas)) +
  geom_point() +
  geom_errorbar(aes(ymin = perTorCas - ci, ymax = perTorCas + ci), width = .25) +
  scale_x_continuous(breaks = seq(1,12), labels = dfM$MonthF, minor_breaks = NULL) +
  scale_y_continuous(limits = c(0, 20)) +
  xlab("") + ylab("Casualty Rate") + 
#  ggtitle(label = "Casualty Rate Per Casualty-Producing Tornado (U.S.)", subtitle = #"1995-2016") +
  theme_minimal()

A = A + ggtitle("A") + 
  theme(plot.title=element_text(hjust=0))
```

Bar graph of casualties by month.
```{r}
ggplot(dfM, aes(x = MonthF, y = nCas)) +
  geom_bar(stat = "identity") +
  scale_y_continuous(limits = c(0, 10000)) +
  xlab("") + ylab("") +
  ggtitle(label = "Number of U.S. Tornado Casualties", subtitle = "1995-2016") +
  theme_minimal()
```

Look at trends in hourly tornado casualties.
```{r}
dfH <- Tor.sfdf %>%
  group_by(hr) %>%
  summarize(nT = n(),
            nCas = sum(cas),
            perTorCas = nCas/nT) %>%
  mutate(ci = qt(0.975, df = nT - 1) * sd(perTorCas)/sqrt(nT))

B = ggplot(dfH, aes(x = hr, y = perTorCas)) +
  geom_point() +
  geom_errorbar(aes(ymin = perTorCas - ci, ymax = perTorCas + ci), width = .25) +
  scale_x_continuous(breaks = seq(0, 24, 4), labels = c("12:00AM", "4:00AM", "8:00AM", "12:00PM", "4:00PM", "8:00PM", "")) +
  scale_y_continuous(limits = c(0, 20)) +
  xlab("") + ylab("Casualty Rate") +
#  ggtitle(label = "Casualty Rate Per Casualty-Producing Tornado (U.S.)", subtitle = #"1995-2016") +
  theme_minimal()

B = B + ggtitle("B") + 
  theme(plot.title=element_text(hjust=0))

source("http://peterhaschke.com/Code/multiplot.R")
#multiplot(A, B, layout = mat)
```

Bar graph of casualties by hr
```{r}
ggplot(dfH, aes(x = hr, y = nCas)) +
  geom_bar(stat = "identity") +
  scale_x_continuous(breaks = seq(0, 24, 4), labels = c("12:00AM", "4:00AM", "8:00AM", "12:00PM", "4:00PM", "8:00PM", "")) +
  scale_y_continuous(limits = c(0, 4500)) +
  xlab("") + ylab("") +
   ggtitle(label = "Number of U.S. Tornado Casualties", subtitle = "1995-2016") +
  theme_minimal()
```

Create a new data frame with columns containing logged and scaled variables.
```{r}
df <- as.data.frame(Tor.sfdf) %>%
  mutate(lpopD = log(popD2),
         lED = log(ED),
         lcas = log(cas),
         smi = scale(MdnIncm),
         smh = scale(MoblHms),
         syr = scale(Year),
         pW = WhitPpl/TotlPpl,
         pB = BlckPpl/TotlPpl,
         plot_date = as.Date(format(Date, "2016-%m-%d")))
```

Year/date plot. Code from: https://buzzfeednews.github.io/2018-07-wildfire-trends/
```{r}
A = ggplot(df, aes(y = Year)) +
  geom_hline(yintercept = seq(1995, 2016, by = 1), color = "gray", size = .05) +
  scale_size_area(max_size = 10, breaks = c(0, 2, 4, 6), 
                  labels = c("1", "7", "55", "403"), name = "Tornado\nCasualties") +
  scale_x_date(date_breaks = "months", date_labels = "%b") +
  scale_y_reverse(limits = c(2016, 1995), breaks = c(2015, 2010, 2005, 2000, 1995)) +
  xlab("") +  ylab("") +
  geom_point(aes(size = lcas, x = plot_date), color = "#fd8d3c", alpha = .5) +
  theme_minimal()
```

Year/hour plot.
```{r}
B = ggplot(df, aes(y = Year)) +
  geom_hline(yintercept = seq(1995, 2016, by = 1), color = "gray", size = .05) +
  scale_size_area(max_size = 10, breaks = c(0, 2, 4, 6), labels = c("1", "7", "55", "403"), name = "Tornado\nCasualties") +
  scale_x_continuous(breaks = seq(0, 24, 4), labels = c("12:00AM", "4:00AM", "8:00AM", "12:00PM", "4:00PM", "8:00PM", "")) +
  scale_y_reverse(limits = c(2016, 1995), breaks = c(2015, 2010, 2005, 2000, 1995)) +
  xlab("") +  ylab("") +
  geom_point(aes(size = lcas, x = hr), color = "#fd8d3c", alpha = .5) +
  theme_minimal()
```

Combine plots
```{r}
source("http://peterhaschke.com/Code/multiplot.R")
mat = matrix(c(1, 2), nrow = 2, byrow = TRUE)
A = A + ggtitle("A") + 
  theme(plot.title=element_text(hjust=0))
B = B + ggtitle("B") + 
  theme(plot.title=element_text(hjust=0))
multiplot(A, B, layout = mat)
```


### Models

Here we extend the model of Elsner et al. (2018) to include a trend term to account for the steady improvement in warning residents that results in a tendency toward fewer casualties. Based on previous research showing a greater risk of casualties for residents in mobile homes we include a term to account for the number of mobile homes estimated within the path.

Try two fixed effects models.
```{r}
library(MASS)
formula1 <- cas ~ syr + lpopD + lED + lpopD:lED + smh
formula2 <- lcas ~ syr + lpopD + lED + lpopD:lED + smh

m1 <- glm.nb(formula1, link = log, data = df)
m2 <- lm(formula2, data = df)

summary(m1)
summary(m2)

cor(df$cas, exp(predict(m1)))
cor(df$cas, exp(predict(m2)))
```

Try some mixed effects models.
```{r}
library(lme4)

formula4 <- lcas ~ syr + lpopD + lED + lpopD:lED + smh + smi + (1|mo) + (1|hr)
m4 <- lmer(formula4, data = df)

summary(m4)
cor(df$cas, exp(predict(m4)))
```

Observed vs predicted.
```{r}
df$pre <- exp(predict(m4))

library(ggrepel)
ggplot(df[df$cas >= 50, ], aes(x = cas, y = pre)) +
  geom_point(aes(size = TotlPpl)) +
  geom_text_repel(aes(label = paste(st, date)), size = 1.75) +
  geom_abline(slope = 1) + 
  geom_smooth(method = lm, se = FALSE) +
  xlab("Observed Number of Casualties") +
  ylab("Predicted Rate") +
  scale_x_log10() + scale_y_log10() +
  theme_minimal()
```

Work with the **brms** package for mixed effects model. First sample from the priors. Set `inits` to `"0"` rather than the default `"random"`.
```{r}
library(brms)

formula5 <- cas | trunc(lb = 1) ~ syr + lpopD + lED + lpopD:lED + smh + (1|hr) + (1|mo)
get_prior(formula5, data = df, family = lognormal)

# Sample from the priors
prior5 <- brm(formula = formula5,
              data = df, family = lognormal,
              prior = c(set_prior("normal(0,5)", class = "b"),
                    set_prior("student_t(3, 3, 10)", class = "Intercept"),
                    set_prior("student_t(3, 0, 10)", class = "sd"),
                    set_prior("student_t(3, 0, 10)", class = "sigma")),
          sample_prior = "only", seed = 9112,
          control = list(adapt_delta = .8))

out <- predict(prior5, probs = c(0, 1))


# Fit the model

post5 <- brm(formula = formula5,
             data = df, family = lognormal,
             prior = c(set_prior("normal(0,5)", class = "b"),
                       set_prior("student_t(3, 3, 10)", class = "Intercept"),
                       set_prior("student_t(3, 0, 10)", class = "sd"),
                       set_prior("student_t(3, 0, 10)", class = "sigma")),
              inits = "0", seed = 9112,
              control = list(adapt_delta = .8))

summary(post5)
```

Correlation between estimated and actual casualties.
```{r}
cor(predict(post5)[, 1], df$cas)
```

NOTE: For different seeds the correlation ranged from .48 to .54. May need to generate more samples.

The `posterior_predict()` function generates 3000 samples of cas. Compare the distribution of cas statistics (mean, max) with actual cas.

```{r}
yrep <- posterior_predict(post5)
df.yrep <- as.data.frame(yrep)
df.out <- reshape2::melt(df.yrep) %>%
  group_by(variable) %>%
  summarize(mx = max(value, na.rm = TRUE),
            mn = exp(mean(log(value), na.rm = TRUE)))

quantile(df.out$mn, probs = c(.25, .75), na.rm = TRUE)
quantile(df.out$mx, probs = c(.25, .75), na.rm = TRUE)
```

Look at the posterior predictive checks.
```{r}
ggplot(df.out, aes(mn)) + 
  geom_density(fill = "red", color = "red") +
  geom_vline(xintercept = exp(mean(log(df$cas))), color = "black", size = 1) +
  scale_x_log10(breaks = c(1, 10, 100, 1000), limits = c(1, 1000)) +
  ylab("Posterior Density") +
  xlab("Average Per-Tornado Casualty Rate") +
  theme_minimal()

ggplot(df.out, aes(mx)) + 
  geom_density(fill = "red", color = "red") +
  geom_vline(xintercept = max(df$cas), color = "black", size = 1) +
  scale_x_log10(breaks = c(100, 1000, 10000, 100000), labels = c(100, 1000, 10000, "100,000")) + 
  ylab("Posterior Density") +
  xlab("Maximum Per-Tornado Casualty Rate") +
  theme_minimal()
```

```{r}
coefTable <- as.data.frame(summary(post5)$fixed) %>%
  mutate(lb = `l-95% CI`,
         ub = `u-95% CI`,
         mi = `Estimate`,
         id = 1:6) %>%
  filter(id %in% c(2, 5, 6)) %>%
  mutate(id2 = 1:3)

ggplot(coefTable, aes(x = id2, y = mi)) +
  geom_hline(yintercept = 0, color = "grey", linetype = "solid") +
  geom_hline(yintercept = c(-.6, -.3, .3), color = "grey", linetype = "dashed") +
  geom_point() +  
  geom_errorbar(aes(x = id2, ymin = lb, ymax = ub), col = "red", width = 0, size = 2) +
  geom_point() +
  scale_x_reverse(breaks = 1:3, labels = c("Trend", "Mobile Homes", "Interaction")) +
  ylab("Fixed Effects") + xlab("") +
  theme_minimal() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  coord_flip() 
```

Save the predicted values and their uncertainties as a data frame. Include additional columns from `Tor.sfdf` for identifying the case (tornado).
```{r}
predValues.df <- as.data.frame(predict(post5)) %>%
  mutate(Actual = df$cas,
         Diff = Actual - Estimate,
         Sign = Diff > 0,
         om = Tor.sfdf$om,
         Year = Tor.sfdf$Year,
         Date = Tor.sfdf$Date,
         st = Tor.sfdf$st,
         slat = Tor.sfdf$slat,
         slon = Tor.sfdf$slon)

predValues.sfdf <- st_sf(predValues.df, 
                         geometry = Tor.sfdf$geometry)

UnderpredValues.df <- predValues.df[predValues.df$Actual >= 25 & predValues.df$Diff >= 0,]

UDTExamples.df <- UnderpredValues.df[c(14, 16, 24, 38, 45, 65, 82),] %>%
  mutate(Names = c("Gainesville, GA", "Spencer, SD", "Camilla, GA", "Camilla, GA", "Marmaduke, AR/Caruthersville, MO", "Smithville, MS/Shottsville, AL", "Garland-Rowlett, TX"))

```

Over/Under prediction plot.
```{r}
library(ggridges)
library(ggrepel)
ggplot(predValues.df[abs(predValues.df$Diff) < 100, ], 
       aes(x = Diff, y = Year, group = Year, height = ..density..)) +
     geom_density_ridges(scale = 3, color = "white") + 
     scale_y_reverse(breaks = seq(1995, 2015, 5)) +
     ylab("") + xlab("Actual - Predicted")

ggplot(predValues.df, aes(y = Year)) +
  geom_hline(yintercept = seq(1995, 2016, by = 1), color = "gray", size = .05) +
  scale_y_reverse(limits = c(2016, 1995), breaks = c(2015, 2010, 2005, 2000, 1995)) +
  xlab("") +  ylab("") +
  geom_point(aes(y = Year, x = Diff), data = predValues.df[predValues.df$Sign, ], color = 'green') +
  geom_point(aes(y = Year, x = Diff), data = predValues.df[!predValues.df$Sign, ], color = 'purple') +
  theme_minimal()

ggplot(predValues.df, aes(x = Actual, y = Estimate)) +
  geom_point() +
  geom_abline(slope = 1) + 
  xlab("Observed Number of Casualties") +
  ylab("Predicted Rate") +
  scale_x_log10() + scale_y_log10() +
  geom_text(data = subset(predValues.df, Actual >= 30),
            aes(label = st), hjust = 1, vjust = .5) +
  theme_minimal()

ggplot(predValues.df[predValues.df$Actual >= 25, ], aes(x = Actual, y = Estimate)) +
  geom_point(col = "#fd8d3c") +
  geom_point(data = UnderpredValues.df, aes(x = Actual, y = Estimate, size = UnderpredValues.df$Diff), col = "#fd8d3c") +
  geom_text_repel(data = UDTExamples.df, aes(label = Names), size = 2.5, nudge_x = .15) +
  geom_abline(slope = 1) + 
#  geom_smooth(method = lm, se = FALSE) +
  xlab("Observed Number of Casualties") +
  ylab("Predicted Rate") +
  scale_x_log10() + scale_y_log10() +
  theme_minimal() +
  theme(legend.position="none")
```

Over/under prediction map.
```{r}
library(USAboundaries)
sts <- state.name[!state.name %in% c("Alaska", "Hawaii")]
stateBorders <- us_states(states = sts)


p1 <- ggplot() +
  geom_sf(data = stateBorders) +
  geom_point(data = predValues.df[predValues.df$Diff >= 0, ], aes(x = slon, y = slat, size = Diff), color = "#1b7837", alpha = .4) +
  scale_size_continuous(name = "Under\nPrediction", 
                        breaks = c(50, 100, 200, 400, 800), 
                        labels = c(50, 100, 200, 400, 800)) +
  xlab("") + ylab("") +
  theme_minimal()

p2 <- ggplot() +
  geom_sf(data = stateBorders) +
  geom_point(data = predValues.df[predValues.df$Diff < 0, ], aes(x = slon, y = slat, size = abs(Diff)), color = "purple", alpha = .3) +
  scale_size_continuous(name = "Over\nPrediction", 
                        breaks = c(50, 100, 500, 1000, 5000), 
                        labels = c(50, 100, 500, 1000, 5000)) +
  xlab("") + ylab("") +
  theme_minimal()

library(patchwork)
p1/p2
```

Using T-Map
```{r}
library(tmap)
tm_shape(stateBorders, projection ="+init=epsg:2163") +
  tm_borders() +
  tm_fill(col = "grey94") +
  tm_shape(predValues.sfdf[predValues.sfdf$Diff >= 0, ]) +
  tm_bubbles("Diff", col = "#fd8d3c", sizes.legend = c(50, 100, 500, 1000, 1500), alpha = 0.9, title.size="Underprediction", scale = 2.5) + 
  tm_shape(stateBorders) + 
  tm_borders(col = "grey", alpha = 0.3) +
  tm_format('World', legend.position = c("left", "bottom"),
                   attr.position = c("left", "bottom"),
                  legend.frame = FALSE) +
  #tm_format_Europe(legend.position = c("left", "bottom"),
  #                 attr.position = c("left", "bottom"),
  #                 legend.frame = TRUE) +
  tm_scale_bar(position = c("right", "bottom")) +
  #tm_compass(position = c("right", "bottom")) +
  tm_layout(frame = FALSE, attr.outside=TRUE)
```

Maps are exported at 8 x 6 in PDF.

### Tornado casualties and place

Top 10 over-predictions of tornado casualties.
```{r}
arrange(predValues.df, Diff)[1:20, ]
```

## Case studies
Camilla, Georgia map
```{r}
library(tmap)
library(tmaptools)
library(albersusa)

states <- usa_sf("longlat")
Region <- states %>%
  filter(iso_3166_2 %in% c("GA", "AL", "MS", "SC", "TN", "NC"))
Region2 <- states %>%
  filter(iso_3166_2 %in% c("GA", "AL", "MS", "SC", "TN", "NC", "FL"))
FL <- states %>%
  filter(iso_3166_2 %in% c("FL"))
counties <- counties_sf("longlat") %>%
  filter(iso_3166_2 %in% c("FL", "GA", "AL"))
Mitchell <- counties %>%
  filter(iso_3166_2 == "GA", name == "Mitchell")
Camilla <- geocode_OSM("Camilla, GA", as.sf = TRUE)

C = tm_shape(states) +
  tm_polygons(col = "grey94") +
tm_shape(Region2, is.master = TRUE) +
  tm_polygons(col = "grey94") +
tm_shape(FL) +
  tm_polygons(col = "grey94") +
#  tm_text(text = "iso_3166_2", xmod = 1.25) +
tm_shape(Region) +
  tm_polygons(col = "grey94") +
#  tm_text(text = "iso_3166_2") +
#tm_shape(Mitchell) +
#  tm_polygons(col = "black") +
#  tm_text(text = "name", ymod = -1, col = "black", size = .75) +
tm_shape(Camilla) +
  tm_dots(col = "#fd8d3c", size = .75, alpha = 0.9) +
#tm_scale_bar(position = c("RIGHT", "BOTTOM")) +
#tm_compass(position = c("right", "bottom")) +
tm_layout(main.title = "Camilla, GA",bg.color="white", attr.outside=TRUE)
```

Spencer, SD map
```{r}
states <- usa_sf("longlat")
Region <- states %>%
  filter(iso_3166_2 %in% c("ND", "SD", "MN", "IA", "NE", "WY"))
MT <- states %>%
  filter(iso_3166_2 %in% c("MT"))
Spencer <- geocode_OSM("Spencer, SD", as.sf = TRUE)

A = tm_shape(states) +
  tm_polygons(col = "grey94") +
tm_shape(Region, is.master = TRUE) +
  tm_polygons(col = "grey94") +
#  tm_text(text = "iso_3166_2") +
tm_shape(MT) +
  tm_polygons(col = "grey94") +
#  tm_text(text = "iso_3166_2", xmod = 1.25) +
tm_shape(Spencer) +
  tm_dots(col = "#fd8d3c", size = .7, alpha = 0.9) +
#tm_scale_bar(position = c("RIGHT", "BOTTOM")) +
#tm_compass(position = c("left", "bottom")) +
tm_layout(main.title = "Spencer, SD",bg.color="white", attr.outside=TRUE)
```

Garland-Rowlett, TX map
```{r}
states <- usa_sf("longlat")
Region <- states %>%
  filter(iso_3166_2 %in% c("TX", "NM", "OK", "AR", "MS"))
LA <- states %>%
  filter(iso_3166_2 %in% c("LA"))
Garland <- geocode_OSM("Garland, TX", as.sf = TRUE)
#Rowlett <- geocode_OSM("Rowlett, TX", as.sf = TRUE)

B = tm_shape(states) +
  tm_polygons(col = "grey94") +
tm_shape(Region, is.master = TRUE) +
  tm_polygons(col = "grey94") +
#  tm_text(text = "iso_3166_2") +
tm_shape(LA) +
  tm_polygons(col = "grey94") +
#  tm_text(text = "iso_3166_2", xmod = -.75) +
tm_shape(Garland) +
  tm_dots(col = "#fd8d3c", size = 2, alpha = 0.9) +
#tm_shape(Rowlett) +
#  tm_dots(col = "red", size = .75, alpha = .75) +
#tm_scale_bar(position = c("RIGHT", "BOTTOM")) +
#tm_compass(position = c("left", "bottom")) +
tm_layout(main.title = "Garland-Rowlett, TX",bg.color="white", attr.outside = TRUE)
```

Smithville, MS/Shottsville, AL map
```{r}
Region <- states %>%
  filter(iso_3166_2 %in% c("GA", "AL", "MS", "SC", "TN", "NC"))
Region2 <- states %>%
  filter(iso_3166_2 %in% c("GA", "AL", "MS", "SC", "TN", "NC", "FL"))
FL <- states %>%
  filter(iso_3166_2 %in% c("FL"))
Shottsville <- geocode_OSM("Shottsville, AL", as.sf = TRUE)

D = tm_shape(states) +
  tm_polygons(col = "grey94") +
tm_shape(Region2, is.master = TRUE) +
  tm_polygons(col = "grey94") +
tm_shape(FL) +
  tm_polygons(col = "grey94") +
#  tm_text(text = "iso_3166_2", xmod = 1.25) +
tm_shape(Region) +
  tm_polygons(col = "grey94") +
#  tm_text(text = "iso_3166_2") +
tm_shape(Shottsville) +
  tm_dots(col = "#fd8d3c", size = .65, alpha = 0.9) +
#tm_scale_bar(position = c("RIGHT", "BOTTOM")) +
#tm_compass(position = c("right", "bottom")) +
tm_layout(main.title = "Smithville, MS/\nShottsville, AL", bg.color="white", attr.outside=TRUE)
```

Gainesville, GA
```{r}
Region <- states %>%
  filter(iso_3166_2 %in% c("GA", "AL", "MS", "SC", "TN", "NC"))
Region2 <- states %>%
  filter(iso_3166_2 %in% c("GA", "AL", "MS", "SC", "TN", "NC", "FL"))
FL <- states %>%
  filter(iso_3166_2 %in% c("FL"))
Gainesville <- geocode_OSM("Gainesville, GA", as.sf = TRUE)

E = tm_shape(states) +
  tm_polygons(col = "grey94") +
tm_shape(Region2, is.master = TRUE) +
  tm_polygons(col = "grey94") +
tm_shape(FL) +
  tm_polygons(col = "grey94") +
#  tm_text(text = "iso_3166_2", xmod = 1.25) +
tm_shape(Region) +
  tm_polygons(col = "grey94") +
#  tm_text(text = "iso_3166_2") +
tm_shape(Gainesville) +
  tm_dots(col = "#fd8d3c", size = .77, alpha = 0.9) +
#tm_scale_bar(position = c("RIGHT", "BOTTOM")) +
#tm_compass(position = c("right", "bottom")) +
tm_layout(main.title = "Gainesville, GA", bg.color="white", attr.outside=TRUE)
```

Marmaduke, AR/Caruthersville, MO
```{r}
Region <- states %>%
  filter(iso_3166_2 %in% c("GA", "AL", "MS", "SC", "TN", "NC"))
Region2 <- states %>%
  filter(iso_3166_2 %in% c("GA", "AL", "MS", "SC", "TN", "NC", "FL"))
FL <- states %>%
  filter(iso_3166_2 %in% c("FL"))
Marmaduke <- geocode_OSM("Marmaduke, AR", as.sf = TRUE)

f = tm_shape(states) +
  tm_polygons(col = "grey94") +
tm_shape(Region2, is.master = TRUE) +
  tm_polygons(col = "grey94") +
tm_shape(FL) +
  tm_polygons(col = "grey94") +
#  tm_text(text = "iso_3166_2", xmod = 1.25) +
tm_shape(Region) +
  tm_polygons(col = "grey94") +
#  tm_text(text = "iso_3166_2") +
tm_shape(Marmaduke) +
  tm_dots(col = "#fd8d3c", size = .55, alpha = 0.9) +
#tm_scale_bar(position = c("RIGHT", "BOTTOM")) +
#tm_compass(position = c("right", "bottom")) +
tm_layout(main.title = "Marmaduke, AR/\nCaruthersville, MO", bg.color="white", attr.outside=TRUE)
```

Plot together
```{r}
tmap_arrange(A, B, C, D, E, f, ncol = 3)
```
